Systematic Design of Optimal Low-Thrust Transfers for the Three-Body Problem \thanks{This research has been supported in part by NSF under the grants CMMI-1243000 (transferred from 1029551), CMMI-1335008, and CNS-1337722.}

### 2017-07-23 Response.tex

\item
    \begin{itshape}
Section 4 ends on page 14 with essentially a thesis statement ("In this paper, we...").  Strive to push the boundary between background and new developments to a much earlier point in the paper; it shouldn't happen so late that you need to remind the audience what the paper is for.
\end{itshape}

Thank you for the comment. 
The introductory detail has been reduced and combined into a single section. 
As a result, the third section begins with the new contributions which describe the optimal control formulation of the reachability set.

\item 
    \begin{itshape}
A missing relevant reference: "On target for Venus — set oriented computation of energy efficient low thrust trajectories" (Dellnitz, Junge, Post, and Thiere, 2006)
\end{itshape}

Thank you for the comment.
The reference has been noted in the discussion of reachability sets.
\begin{correction}
More recently, reachability theory has recently been applied to space systems~\cite{holzinger2009,komendera2012a,dellnitz2006}.
\end{correction}

\item
    {\color{red}
    \begin{itshape}
*** Variational Integrators
Many of the specific claims regarding the use of a variational integrator are unsubstantiated and, I believe, incorrect. Why is energy drift "even more detrimental" with regards to "the long term effects of low-thrust"?  I would need to be convinced that a non-geometric integrator's energy drift relative to a constant energy baseline is significantly different from its drift relative to a gradually shifting baseline corresponding to low thrust. The specific claims about the numerical stability of derivatives also don't seem related to energy preservation.

\end{itshape}
}

\item
    {\color{red}
    \begin{itshape}
Apart from these claims, I question the need for a variational integrator in the first place. The example in section 3.4 is not convincing because there is no apples-to-apples basis for the comparison; a variable-step integrator with unspecified tolerance is compared to a fixed-step integrator with an arbitrary step size.  Integrator comparisons should consider error in proportion to compute cost, which can be tuned by step size or tolerance.  Also, fixed step seems inappropriate for a path through diverse dynamic regimes, and it's not clear that the problem scope calls specifically for energy preservation (especially since the idea is to use low thrust to mitigate some time cost of exploiting the problem structure).  
\end{itshape}
}

@@

\item
    \begin{itshape}
*** Figures
Altogether, the figures need more descriptive captions.  
\end{itshape}

Thank you for the comment, and we agree that the figures will benefit from more descriptive captions.
We've added additional descriptions for all of the figures.
The figure captions are shown below, with the figures reduced in size to save some space.

\begin{correction}
\begin{figure}[H]
        \centering
        \includegraphics[width=0.2\textwidth]{energy_contours}
        \caption{Contour plot of Jacobi integral: Zero velocity curves of constant Jacobi integral. A particle with fixed energy level cannot cross the contour lines and is therefore limited to specific regions of the phase space. }
\end{figure}

\begin{figure}[H]
        \centering
        \begin{scaletikzpicturetowidth}{0.2\textwidth}
    \begin{tikzpicture}[tdplot_main_coords,
      poincare/.style={opacity=.2,very thick,fill=blue},
      orbit/.style={very thick,black},
      orbit hidden/.style={very thick,dashed},
      grid/.style={very thin,gray!50},
      axis/.style={->,blue,thick}, scale=\tikzscale]

@@

    \tdplotdrawarc[orbit hidden]{(center)}{2}{90}{190}{}{};
    \tdplotdrawarc[orbit,<-]{(center)}{2}{-170}{90}{}{};

    \tdplotdrawarc[orbit hidden]{(center)}{3}{90}{199}{}{};
    \tdplotdrawarc[orbit,<-]{(center)}{3}{-161}{90}{}{};

        \end{tikzpicture}
    \end{scaletikzpicturetowidth}
    \caption{Diagram of the \Poincare map: Periodic orbits will appear as fixed points on the \Poincare section \( \Sigma \). Stabiilty of periodic orbits is clearly visible on the section as successive intersections approach or depart a fixed point.}
\end{figure}

\begin{figure}[H]
        \centering 
        \subcaptionbox{Earth-Moon three-body trajectory used for integrator comparision: A stable trajectory is used to test the variational and conventional integrators. 
        The energy level is high enough to enter the vicinity of the Moon but not escape the three body system.}
        [0.5\textwidth]{\includegraphics[width=0.2\textwidth, height=0.2\textheight, keepaspectratio]{trajectory}}~
        \subcaptionbox{Mean Jacobi integral deviation between variational integrator and Matlab \texttt{ODE45}: A range of fixed step sizes are used for the variational integrator to approximately match the computational time of \texttt{ODE45}}
        [0.5\textwidth]{\includegraphics[width=0.2\textwidth, height=0.2\textheight, keepaspectratio]{figures/cputimevsE.pdf}}~
\end{figure}

\begin{figure}[H]
        \centering
%       \includegraphics[width=0.5\textwidth]{reachability_set_hand}
        \begin{scaletikzpicturetowidth}{0.2\textwidth}
        \begin{tikzpicture}[tdplot_main_coords,
          poincare/.style={opacity=.2,very thick,fill=blue},
          orbit/.style={very thick,black},
          orbit hidden/.style={very thick,dashed},
          grid/.style={very thin,black},
          axis/.style={->,blue,thick},
          reachability/.style={thick,blue},scale=\tikzscale]

@@

                    % draw markers at each subsegment

                    % draw the subsegments
                    \draw [->] (x0.center) to [bend left=30] (x1minus.center);
                    \draw [->] (x1plus.center) to [bend left=30] (x2minus.center);
                    % dotted lines here
                    \draw[decorate sep={1mm}{5mm},fill] (x2plus.east) to [bend left=10] (x3minus.west);

                    \draw [->] (x3plus.center) to [bend left=30] (xnminus.center);
                \end{tikzpicture}
        \end{scaletikzpicturetowidth}
        \caption{Schematic diagram of the multiple shooting method:
        The complete trajectory is split into a nubmer of sub-segments, and additional interior constraints are included to ensure state and costate continunity.
    Splitting the optimal trajectory into short segments reduces the sensitivity of terminal states to variations of the initial states.}
\end{figure}

\begin{figure}[H]
   \centering
   \includegraphics[width=0.2\textwidth]{moon_orbit} % requires the graphicx package
   \caption{\(L_1\) periodic orbit transfer to orbit of the Moon: Example scenario demonstraiting the initial and target orbit.
   Without the low-thrust proplusion, the spacecraft is constrained to the initial periodic orbit. 
   We determine the reachability set to find a transfer trajectory to the target orbit about the moon.}
\end{figure}

\begin{figure}[H]
        \centering 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=0.5\textwidth]{manifold_trajectory} 
                \caption{Position space view of invariant manifold} \label{fig:manifold_trajectory} 
        \end{subfigure}~ %
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=0.5\textwidth]{manifold_poincare} 
                \caption{\Poincare intersection of invariant manifold (green), target orbit (blue) and the initial orbit (black)} \label{fig:manifold_poincare} 
        \end{subfigure} 
        \caption{Invariant manifold transfer: An example transfer using the invariant manifolds is shown in both the position and \Poincare spaces.
        The control free transfer from the initial periodic orbit to the target orbit result in a long time of flight. 
    In addition, the manifold only intersects the target orbit on the descending or far side of the moon.}
\end{figure}
\begin{figure}[H]
        \centering 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=0.5\textwidth]{reach_trajectory} 
                \caption{Position space view of reachability set}  
        \end{subfigure}~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line) 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=0.5\textwidth]{poincare_compare} 
                \caption{\Poincare section view of reachability set}  
        \end{subfigure} %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. %(or a blank line to force the subfigure onto a new line)

        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=0.5\textwidth]{reach_transfer} 
                \caption{Transfer trajectory selected from reachabiltiy set viewed in the position space}  
        \end{subfigure}~ 
        \begin{subfigure}[htbp]{0.5\textwidth} 
                \includegraphics[width=0.5\textwidth]{control_input_l1} 
                \caption{Control input for the selected transfer trajectory}  
        \end{subfigure}~
        \caption{\( L_1 \) Reachability set transfer: The low thrust propulsion is used to approximate the reachability set starting from the initial periodic orbit over a fixed time horizon.
        The reachability set is shown in the upper two figures in both the position and \Poincare space.
    From this reachabiilty set we chose a trajectory which intersects the target orbit and it is shown in the lower left figure.
The optimal control to achieve this transfer is shown in the lower right figure.} 
\end{figure}

\begin{figure}[H]
   \centering
   \includegraphics[width=0.2\textwidth]{initial_final} % requires the graphicx package
   \caption{Geostationary to \( L_1 \) transfer: Representation of the initial and target orbits for the reachability transfer. 
   Vehicle is assumed to begin on the geostationary orbit about \( m_1\) and will transfer to the periodic orbit about \( L_1\)}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage1_trajectory_zoom.pdf} 
        \caption{Stage 1 Trajectory} 
    \end{subfigure}~
    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage1_poincare.pdf} 
        \caption{Stage 1 \Poincare~section } 
    \end{subfigure}

    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage2_trajectory_zoom.pdf} 
        \caption{Stage 2 Trajectory~} 
    \end{subfigure}~
    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage2_poincare.pdf} 
        \caption{Stage 2 \Poincare~section } 
    \end{subfigure}

    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage3_trajectory_zoom.pdf} 
        \caption{Stage 3 Trajectory} 
    \end{subfigure}~
    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage3_poincare.pdf} 
        \caption{Stage 3 \Poincare~section } 
    \end{subfigure}

    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_trajectory_zoom.pdf} 
        \caption{Stage 4 Trajectory~} 
    \end{subfigure}~
    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage4_poincare.pdf} 
        \caption{Stage 4 \Poincare~section } 
    \end{subfigure}   
    \caption{Stage 1-4 reachability sets: The first four reachability sets visualized in both the position (left) and \Poincare space (right).
        The minimum trajectories from the preceeding stages are shown in red, while the next stage is shown in blue.
    The transfer goal is to generate a complete trajectory from the initial geostationary orbit to the \( L_1 \) stable manifold, with an eventual arrival at the \( L_1 \) periodic orbit}
\end{figure}

\begin{figure}[H]
    \centering
    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_trajectory_zoom.pdf} 
        \caption{Stage 5 Trajectory~} 
    \end{subfigure}~
    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage5_poincare.pdf} 
        \caption{Stage 5 \Poincare~section } 
    \end{subfigure}

    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage6_trajectory_zoom.pdf} 
        \caption{Stage 6 Trajectory} 
    \end{subfigure}~
    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage6_poincare.pdf} 
        \caption{Stage 6 \Poincare~section } 
    \end{subfigure}

    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage7_trajectory_zoom.pdf} 
        \caption{Stage 7 Trajectory} 
    \end{subfigure}~
    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage7_poincare.pdf} 
        \caption{Stage 7 \Poincare~section } 
    \end{subfigure}

    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage8_trajectory_zoom.pdf} 
        \caption{Stage 8 Trajectory} 
    \end{subfigure}~
    \begin{subfigure}[htbp]{0.2\textwidth} 
        \includegraphics[width=\textwidth, keepaspectratio]{figures/geo_transfer/stage8_poincare.pdf} 
        \caption{Stage 8 \Poincare~section } 
    \end{subfigure}   
    \caption{Stage 5-8 reachability sets: The last four reachability sets visualized in both the position (left) and \Poincare space (right).
        The minimum trajectories from the preceeding stages are shown in red, while the next stage is shown in blue.
    The transfer goal is to generate a complete trajectory from the initial geostationary orbit to the \( L_1 \) stable manifold, with an eventual arrival at the \( L_1 \) periodic orbit}
\end{figure}
\end{correction}

\item
    \begin{itshape}
Figure 4 does not show what I think you intend it to show; it looks like two periodic circular orbits.
\end{itshape}

Thank you for the comment, however the figure is drawn to show a pictoral representation of a \Poincare section.
The periodic orbits intersect the section at fixed points, as a result it becomes simple to visualize and determine these fixed solutions by searching for fixed points on the section
Conversely, stable/unstable trajectories will also appear on the section with a distinctive signature, and by tracking successive intersections of the trajectory with the \Poincare section.

In order to improve the clarity and reduce confusion, we have added some additional detail to the discussion of the \Poincare map to the manuscript.
This addition has been copied below.

\begin{correction}
\Poincare maps are a useful tool in the analysis of the flow near periodic orbits in the three-body problem.
We let \( \Sigma \) define a hypersurface of section chosen such that all trajectories in the vicinity of a state \( \vecbf{q} \in \Sigma \) cross \( \Sigma \) transversely and in the same direction.
A \Poincare map, \( P(\vecbf{q}) = \phi(T;\vecbf{q}) \), maps the state of a trajectory from one intersection to the next.
Choosing a section in this manner results in a \Poincare section as shown in~\cref{fig:poincare_map}.
In~\cref{fig:poincare_map} we show two examples of periodic trajectories intersecting the \Poincare section. 
Periodic solutions will appear as fixed points on the section, such as \( q_0, q_1 \) in~\cref{fig:poincare_map}, while stable or unstable trajectories become clearly visible by viewing successive intersections of the section.
This allows for greater insight into the stability and dynamics of periodic solutions of a dynamic system as a fixed point on the \Poincare section corresponds to a periodic orbit while movement on the section is associate with the stability of neighboring trajectories. 
\end{correction}

### 2017-07-23 Response.tex

Thank you for the comment.
Appendix B has been removed from the revised manuscript.
The manuscript has been modified to mention that Gauss Jordan elimination is used to avoid the use of the matrix inversion.

\begin{correction}
The derivation of~\cref{eq:costate_eom} is given in Appendix A.
In addition, the computation of~\cref{eq:costate_eom} requires inversion of the Jacobian matrix.
This is a computationally expensive operation that is prone to numerical error and instability.
We use Gauss-Jordan elimination to avoid this inversion in~\cref{eq:costate_eom} and determine an explicit update map \( \vecbf{\lambda}_k \to \vecbf{\lambda}_{k+1} \).
\end{correction}

\item 
    \begin{itshape}
In line 14 (page 12), you define the state transition matrix over one orbital period, and then on line 18 you start referring to it as the monodromy matrix.  I understand what you're doing, but you may want to make the definition more explicit.  Some of this section can be substantially reduced and other source material could be referenced anyway.
\end{itshape}

Thank you for the comment.
The discussion about the invariant manifolds and \Poincare sections has been reduced.
This discussion has been removed from the revised manuscript and replaced with a general summary and some relevant citations.
This specific discussion no longer exists in the manuscript but rather cites several references with the details on determining the invariant manifolds.

@@

\item 
    \begin{itshape}
It may also be generally worth acknowledging the benefits of Taylor's method within the context of variational integrators:
Schmitt, J., Shingel, T., Leok, M., "LAGRANGIAN AND HAMILTONIAN TAYLOR VARIATIONAL INTEGRATORS," eprint arXiv:1703.06599, 2017

Online copy: https://arxiv.org/pdf/1703.06599.pdf 
\end{itshape}

Thank you for the comment.
However, this manuscript is not really focused on variational integrators but rather the design of transfer trajectories using reachability sets.
There is a wide range of topic under the umbrella of varaitional integrators that are not directly applicable to this manuscript and work.
Interested readers can find a wide variety of literature on the relative benefits of all manner of variational integrators in papers focused on that topic specifically.
As a result, we do not feel it is applicable to include additional discussion on the merits of taylor integration methods.

@@

Thank you for the comment.
The additional clarification has been added to the manuscript and copied below.
\begin{correction}

Over long simulation horizons or with the addition of small control inputs the inability of conventional integration schemes to accurately track the system energy limits the applicability of conventional techniques in which energy conservation is mandatory for characterizing the solution space.
\end{correction}

\item 
    \begin{itshape}
Section 5 - This method, while very useful in its given context is highly dependent on the use of the PCRTB in the framework of the current analysis.  This needs to somehow be conveyed.  3D transfers are significantly more challenging when it comes to exploiting Poincare sections.  It can be done, but more subplots, dimensions, and data storing are necessary to view the subspace - save to say, it is a more challenging endeavor that must be acknowledged for general purpose low-thrust mission design.
\end{itshape}

An additional paragraph has been added to the beginning of the section which discusses our reachability based transfer approach.
This additional content is copied below.

@@

Additional detail has been incorporated which discusses the specific solver used in computing the solutions. 
The description does not specifically mention all of the possible options in \texttt{fsolve} but rather gives the key components used in our solution.
This approach seems to be consistent with the literature, e.g.~\cite{ozimek2010a,stuart2010}.
The modified portion is copied below.

\begin{correction}
In this work, we use the Matlab nonlinear solver \texttt{fsolve} to solve the system of nonlinear equations defined by the multiple shooting algorithm.
Within \texttt{fsolve}, we use the trust-region dogleg solver which makes use of the Powell dogleg procedure for computing a step direction and magnitude to minimize successive iterations of the solver.
All numerical integration is performed using the discrete variational integrator described in~\cref{sec:discrete_var}.
\end{correction}

\item
    \begin{itshape}
Did the authors experiment with the number of multiple shooting stages in their example problems?  There must be some sort of "sweet spot" in number of stages for the example problems attempted.  If I missed it, I apologize, but I didn't see it.
\end{itshape}

We did spend considerable time in determining the appropriate number of stages. 
One key issue was that our implementation in software depended on dividing the total number of integration steps by the number of interior stages.
For simplicity, this reduced to possible number of stages to an even number. 
Additionally, we discovered that adding many stages tended to degrade the solution and cause difficulties in covergence. 
The solver would tend to ensure a continuous trajectory, by satisfying the interior stages, at the expense of the terminal stage. 
Some additional detail has been provided that mentions our choice of the number of stages and has been copied below.

@@

\begin{correction}
Based on experimentation, we use four stages in our multiple shooting method.
This provided the best performance and convergence stability while minimizing the difficulties in additional interior constraints.
\end{correction}

\item 
    \begin{itshape}
Section 6.1 - I don't think you need to explain how you created the three-body orbits.  This can be found in many references - You can just cite a reference and give the orbit specifications.  (Starting with line 49, page 18, up to line 21 page 19.)
\end{itshape}

The discussion of generating periodic orbits in the three body problem has been removed.
It replaced with a small introduction mentioning the differential correction process and a citation to a relevant resource. 
The revised paragraph is copied below.

@@

\item
    \begin{itshape}
"Low energy transfers from the Earth to the Lagrange points are necessary for future missions."  One must be careful not to confuse these with traditional "low-energy transfers" to the moon, which use Sun third-body perturbations to transfer into lunar orbit.  These are what many mission designers define as "low-energy" lunar transfers.  This is actually a four-body problem (Sun, Earth, moon, spacecraft), and represents additional delta-V savings (as was done with NASA's GRAIL mission.)

Example:

Parker, J. and Anderson, R., "Low-Energy Lunar Trajectory Design," DEEP SPACE COMMUNICATIONS AND NAVIGATION SERIES, Jet Propulsion Laboratory Pasadena, California, July 2013.

Online Copy: https://descanso.jpl.nasa.gov/monograph/series12/LunarTraj--Overall.pdf
\end{itshape}
This sentence has been modified to make it more clear that we're discussing low thrust propulsion rather than low-energy transfers.
The modified sentence is copied below.

@@

\begin{correction}
Transfers from the Earth to the Lagrange points, through the use of low-thrust electric propulsion, offer an additional and potentially shorter time of flight in comparison to the low-energy transfers which utilize solar perturbations.
\end{correction}

\item 
    \begin{itshape}
For what it's worth, exploiting sun perturbations with low-thrust in the Earth-moon RTBP would be highly beneficial and represent a worthy avenue of future research (more on this below.)
\end{itshape}

This is a very useful and interesting suggestion and something that we will take into consideration for future research.
